Kelp morphometric properties and the link to hydrodynamic modelling

### 2018-09-20 chapter_1.Rmd
edited and added to the wave exposure section.

---
title: "Kelp morphometric properties and the link to hydrodynamic modelling"
author: "Ross Coppin"
date: "26 February 2017"
output:
  word_document:
    toc: yes
  pdf_document:
    highlight: zenburn
    number_sections: yes
    toc: yes
---

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception. Wave exposure has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of macroalgae communties. For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure. Wave exposure may also drive macroalgae communties indirectly through the alteration in affect of another environmental driver. For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundry layer around the frond [@lobban1994]. The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgment, which ultimately leads to expiration. Wave exposure is a complex abiotic variable which varies spatially and temporarly in the marine enviroment. Furthermore, the degree to which a macroalgae community is exposed which depend on local site characteristics, such as bathymetry and local wind patterns. Despite this fact, macrolagae have been able to persist in often harsh and variable wave environments. Macroalage are sessile organisms and incapabable of migrating when local conditions become unsuitable. Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological variation.  @moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid concentrations.

>>>>

---
title: "Kelp morphometric properties and the link to hydrodynamic modelling"
author: "Ross Coppin"
date: "26 February 2017"
output:
  pdf_document:
    highlight: zenburn
    number_sections: yes
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
---

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception. Wave exposure has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of macroalgae communties. For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure. Wave exposure may also drive macroalgae communties indirectly through the alteration in affect of another environmental driver. For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundry layer around the frond [@lobban1994]. The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgment, which ultimately leads to expiration. Wave exposure is a complex abiotic variable which varies spatially and temporarly in the marine enviroment. Furthermore, the degree to which a macroalgae community is exposed which depend on local site characteristics, such as bathymetry and local wind patterns. Despite this fact, macrolagae have been able to persist in often harsh and variable wave environments. Macroalage are sessile organisms and incapabable of migrating when local conditions become unsuitable. Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological variation. This morphological variation allows macroalgae to reduce the amount of drag experienced during times of high wave energy or pulse distrubance events, such as storms. This may be achieved through different strategies and are species-specific, which can be directly attributed the high diversity in morphological characters of algal communities. For example some algae have fronds and others are articulated corraline, and therefore these species would need to adpat their morphology differently in order to persist. @moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid concentrations.

### 2018-09-20 chapter_1.tex
edited and added to the wave exposure section.

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{eurosym}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Kelp morphometric properties and the link to hydrodynamic modelling},
            pdfauthor={Ross Coppin},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

The important environmental drivers of kelp idividuals and communities
include light, substrata, salinity, sedimentation, nitrients,
temperature and water motion. Although studies have investigated the
effects of important environmental drivers, the roles these factors play
is often difficult to evaluate as such factors may never be fully
independent of each other, i.e.~environmental factors are to some extent
temperature dependent. Multifactorial studies have attempt to explain
combined affects, however these studies are often limited to
investigating combination of two or three environmental drivers as
inclusion of too many factors can lead to results that are difficult to
interept. Environmetal factors are highly variable on temporal and
spatial scales, and their effects may also dependent on life-stage
adding a further layer of complexity to investigations.

@@

Light is an important factor for kelp survival, however if light is
limited or excessive this may negatively impact kelp survival or growth.
Much of the past research into the role light plays into the functioning
of kelp {[}@Bruhn1996);@Belseth2012{]}. For instance, solar ultraviolate
radiation has been shown to aï¬\euro{}ect sub-canopy Ecklonia radiata
sporophytes when the canopy of mature \emph{Ecklonia radiata} was
removed {[}@Wood 1987{]}. The sub-canopy sporophytes experienced tissue
damage,photopigment destruction,reduced growthand decreased
survivorship, thus inhibiting their settlement and survival {[}@Wood
1987{]}. Laboratory experiments revealed that the UV component of
radiation, rather than intense radiation itself, was responsible for the
eï¬\euro{}ects mentioned above. High light stress has negative
eï¬\euro{}ects, such as photoinhibition and photo-damage on
\emph{Ecklonia cava} sporophytes {[}@Altamirano2004{]}. @Altamirano2004
found that \emph{Ecklonia cava} is more vulnerable to light stress
conditions, and less likely to recover under unfavourable conditions
{[}@Altamirano2004{]}. @Bolton1985 showed that under sub-saturating
irradiances and supra- optimal temperatures \emph{Ecklonia maxima} to
showed a decrease in reproductive rates and an increase in cell
production. An additional ï¬nding of this study was that despite the
decrease in reproductive rates, the ï¬nal egge production per female
was greater under these conditons. The authors interpreted this an
ecological adaption that may increase survival rates under times of
stress or non - ideal conditions {[}@Bolton1985{]}.

@@

Depth does not aï¬\euro{}ect kelp ecosystems directly, however a change
in depth often causes ï¬‚uctuations or changes in other environmental
variables such as water motion, light and temperature. Water motion also
decreases with depth, and some kelps better suited to deeper
environments (\emph{L. pallida}) replace those in the shallows (\emph{E.
maxima}) (Dayton 1985; Gerard 1982). The increase in depth can lead to a
decrease in sunlight penetration, with some species better adapted for
low-light conditions than others, such as (\emph{L. pallida}).
Temperature may also change along a depth gradient due to a reduction in
sunlight penetration. (Dayton 1985; Gerard 1982). Therefore depth does
not directly play a role in kelp functioning but may alter more
inï¬‚uential factors such as light and water motion.

@@

The importance of nutrients in the functioning of kelps is well
understood {[}@Dayton1985; @Gaylord2012{]}. Dissolved nitrogen, and in
particular nitrate, are important; however research has also placed
emphasis on phosphate and other trace compounds for functioning of kelps
{[}@Dayton1985{]}. Additionally, some kelps have the ability to store
inorganic nitrogen in order to compensate for periods of low nutrient
availability, which has been observed for Laminaria and Macrocystis
{[}@Dayton1985; @Gaylord2012{]}. Nutrient stratiï¬cation is also an
important factor, particularly for canopy type kelps. The concentration
of nutrients at the surface is important to the functioning and
maintenance of the canopy. For instance kelp canopies in California
often deteriorate in the summer months when surface nitrate levels are
low {[}@Jackson1977{]}. Water motion is important in the assimilation of
nutrients from the water column, and kelps have been shown to adapt
blade morphology in order to create more turbulence around the boundary
layer of the frond to enhance nutrient assimilation (Wheeler 1980).
Temperature has also been closely linked with nutrient concentrations.
Nutrients are often in higher concentrations in the water column during
low temperature events. This is often an indication of an
â\euro{}œup-wellingâ\euro{} event, which brings cold and nutrient rich
waters from the bottom to the surface of the water column. Temperature
can play a direct role in the uptake of nutrients through eï¬\euro{}ects
on algal metabolism; however this may vary from species to species
{[}@Raven1988{]}.

@@

Temperature is a driver of kelp species distributions and
ecophysiological processes, as well as a lesser role in morphological
adaption\ldots{}\textbf{example here}\ldots{}The majority of kelp
species are artic and temperate organisms, and the warming of ocean
temperatures is expected to cause a poleward biogeographical shift of
species {[}@Bolton2012{]}. There is evidence to suggest that South
African kelp forests are expanding due to ocean cooling
{[}@Bolton2012{]}, possibly driven by an intensiï¬cation and increase
in coastal upwelling {[}@Blamey2012, @Blamey2015{]}. In South Africa
there has been a biogeographical shift eastward along the coast due to a
change in inshore temperature regime, making South Africa no exception
to changing ocean temperatures {[}@Bolton2012{]}. Macroalgae, such as
kelps, can react to an increase in surface temperatures in one of three
ways: they can migrate, adapt and die {[}@Biskup2014{]}. A study by
@Biskup2014 investigated the functional response of two kelp species
(\emph{Laminaria ochroleuca} and \emph{Saccorhiza polyschides}) to
rising sea temperatures. The functional responses of Saccorhiza
polyschides was measured for both the subtidal and intertidal habitats,
to see what aï¬\euro{}ect non- optimal conditions (intertidal zone) had
on the kelps {[}@Rinde2005{]}. The study found that Laminaria ochroleuca
exhibited a poor ability to acclimatise and was dependent on the
kelpâ\euro{}™s life history traits {[}@Biskup2014{]}. Therefore annual
kelp species are more likely to survive under non-ideal condition, and
the intertidal Saccorhiza polyschides, compared to the subtidal, showed
a higher physiological ï¬‚exibility to changing conditions
{[}@Biskup2014{]}. This may be because the intertidal zone undergoes far
more change than the subtidal and therefore kelps in the intertidal are
forced to adapt to harsher conditions where ï¬‚uctuations in
temperature, sunlight, turbidity and water motion are common. The
eï¬\euro{}ects on temperature have also been investigated by
@Wernberg2010. The study looked at resilience of kelp beds along a
latitudinal temperature gradient. Kelp abundance is likely to decline
with the predicted warming of ocean waters @Wernberg2010 and although
kelps have the ability to acclimatize and adjust their metabolic
performance, which in turn allows them to change their physiological
performance to mitigate the seasonal ï¬‚uctuations in temperature, this
acclimatization is done at a cost @Wernberg2010\ldots{}\textbf{link to
paragraph on kelp morphology}\ldots{}

@@

Other than temperature, wave exposure is also recognised as an important
driver of the marine environment. Like temperature, wave exposure can
act on regional scales and has been shown to play a role in determing
distribution, abundance, diversity, composition and productivity of
benthic and rocky shore communties. For example, increasing degrees of
exposure may positively influence the amount of area available to trap
light on macroalgal fronds, as well as increasing nutrient uptake
through increased turbulence in the boundry layer {[}@lobban1994{]}.
@lewis found that degrees of wave exposure vertical structure rocky
shore communities\ldots{}\textbf{work in progress}\ldots{}

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Kelp morphometric properties and the link to hydrodynamic modelling},
            pdfauthor={Ross Coppin},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

The important environmental drivers of kelp individuals and communities
include light, substrata, salinity, sedimentation, nitrients,
temperature and wave exposure. Although studies have investigated the
effects of important environmental drivers, the roles these factors play
is often difficult to evaluate as such factors may never be fully
independent of each other, i.e.~environmental factors are to some extent
dependent on one another. Multifactorial studies have attempted to
explain combined affects, however these studies are often limited to
investigating combination of two or three environmental drivers as
inclusion of too many factors can lead to results that are difficult to
interept. Environmetal factors are highly variable on temporal and
spatial scales and their effects may also be dependent on the life-stage
of the organism, adding a further layer of complexity to investigations.

@@

Light is an important factor for kelp survival, however if light is
limited or excessive this may negatively impact kelp survival or growth.
Much of the past research into the role light plays into the functioning
of kelp {[}@Bruhn1996);@Belseth2012{]}. For instance, solar ultraviolate
radiation has been shown to aﬀect sub-canopy Ecklonia radiata
sporophytes when the canopy of mature \emph{Ecklonia radiata} was
removed {[}@Wood 1987{]}. The sub-canopy sporophytes experienced tissue
damage,photopigment destruction,reduced growthand decreased
survivorship, thus inhibiting their settlement and survival {[}@Wood
1987{]}. Laboratory experiments revealed that the UV component of
radiation, rather than intense radiation itself, was responsible for the
eﬀects mentioned above. High light stress has negative eﬀects, such as
photoinhibition and photo-damage on \emph{Ecklonia cava} sporophytes
{[}@Altamirano2004{]}. @Altamirano2004 found that \emph{Ecklonia cava}
is more vulnerable to light stress conditions, and less likely to
recover under unfavourable conditions {[}@Altamirano2004{]}. @Bolton1985
showed that under sub-saturating irradiances and supra- optimal
temperatures \emph{Ecklonia maxima} to showed a decrease in reproductive
rates and an increase in cell production. An additional ﬁnding of this
study was that despite the decrease in reproductive rates, the ﬁnal egge
production per female was greater under these conditons. The authors
interpreted this an ecological adaption that may increase survival rates
under times of stress or non - ideal conditions {[}@Bolton1985{]}.

\section{Substrata}\label{substrata}

\section{Salinity}\label{salinity}

@@

Depth does not aﬀect kelp ecosystems directly, however a change in depth
often causes ﬂuctuations or changes in other environmental variables
such as water motion, light and temperature. Water motion also decreases
with depth, and some kelps better suited to deeper environments
(\emph{L. pallida}) replace those in the shallows (\emph{E. maxima})
{[}@Dayton1985; @Gerard1982{]}. The increase in depth can lead to a
decrease in sunlight penetration, with some species better adapted for
low-light conditions than others, such as (\emph{L. pallida}).
Temperature may also change along a depth gradient due to a reduction in
sunlight penetration {[}@Dayton1985; @Gerard1982{]}. Therefore depth
does not directly play a role in kelp functioning but may alter more
inﬂuential factors such as light and water motion.

\section{Sedimentation}\label{sedimentation}

@@

The importance of nutrients in the functioning of kelps is well
understood {[}@Dayton1985; @Gaylord2012{]}. Dissolved nitrogen, and in
particular nitrate, are important; however research has also placed
emphasis on phosphate and other trace compounds for functioning of kelps
{[}@Dayton1985{]}. Additionally, some kelps have the ability to store
inorganic nitrogen in order to compensate for periods of low nutrient
availability, which has been observed for Laminaria and Macrocystis
{[}@Dayton1985; @Gaylord2012{]}. Nutrient stratiﬁcation is also an
important factor, particularly for canopy type kelps. The concentration
of nutrients at the surface is important to the functioning and
maintenance of the canopy. For instance kelp canopies in California
often deteriorate in the summer months when surface nitrate levels are
low {[}@Jackson1977{]}. Water motion is important in the assimilation of
nutrients from the water column, and kelps have been shown to adapt
blade morphology in order to create more turbulence around the boundary
layer of the frond to enhance nutrient assimilation (Wheeler 1980).
Temperature has also been closely linked with nutrient concentrations.
Nutrients are often in higher concentrations in the water column during
low temperature events. This is often an indication of an ``up-welling''
event, which brings cold and nutrient rich waters from the bottom to the
surface of the water column. Temperature can play a direct role in the
uptake of nutrients through eﬀects on algal metabolism; however this may
vary from species to species {[}@Raven1988{]}.

@@

Temperature is a driver of kelp species distributions and
ecophysiological processes, as well as a lesser role in morphological
adaption\ldots{}\textbf{example here}\ldots{}The majority of kelp
species are artic and temperate organisms, and the warming of ocean
temperatures is expected to cause a poleward biogeographical shift of
species {[}@Bolton2012{]}. There is evidence to suggest that South
African kelp forests are expanding due to ocean cooling
{[}@Bolton2012{]}, possibly driven by an intensiﬁcation and increase in
coastal upwelling {[}@Blamey2012, @Blamey2015{]}. In South Africa there
has been a biogeographical shift eastward along the coast due to a
change in inshore temperature regime, making South Africa no exception
to changing ocean temperatures {[}@Bolton2012{]}. Macroalgae, such as
kelps, can react to an increase in surface temperatures in one of three
ways: they can migrate, adapt and die {[}@Biskup2014{]}. A study by
@Biskup2014 investigated the functional response of two kelp species
(\emph{Laminaria ochroleuca} and \emph{Saccorhiza polyschides}) to
rising sea temperatures. The functional responses of Saccorhiza
polyschides was measured for both the subtidal and intertidal habitats,
to see what aﬀect non- optimal conditions (intertidal zone) had on the
kelps {[}@Rinde2005{]}. The study found that Laminaria ochroleuca
exhibited a poor ability to acclimatise and was dependent on the kelp's
life history traits {[}@Biskup2014{]}. Therefore annual kelp species are
more likely to survive under non-ideal condition, and the intertidal
Saccorhiza polyschides, compared to the subtidal, showed a higher
physiological ﬂexibility to changing conditions {[}@Biskup2014{]}. This
may be because the intertidal zone undergoes far more change than the
subtidal and therefore kelps in the intertidal are forced to adapt to
harsher conditions where ﬂuctuations in temperature, sunlight, turbidity
and water motion are common. The eﬀects on temperature have also been
investigated by @Wernberg2010. The study looked at resilience of kelp
beds along a latitudinal temperature gradient. Kelp abundance is likely
to decline with the predicted warming of ocean waters @Wernberg2010 and
although kelps have the ability to acclimatize and adjust their
metabolic performance, which in turn allows them to change their
physiological performance to mitigate the seasonal ﬂuctuations in
temperature, this acclimatization is done at a cost
@Wernberg2010\ldots{}\textbf{link to paragraph on kelp
morphology}\ldots{}

@@

Other than temperature, wave exposure is also recognised as an important
driver of the marine environment, and macroalgae are not exception. Wave
exposure has been shown to play a role in determing distribution,
abundance, diversity, composition, growth {[}@Cousens1982{]} and
productivity {[}@pedersen2012a{]} of macroalgae communties. For example,
the width, vertical zonation and diversity of algal communities often
change predictably along gradients of wave exposure. Wave exposure may
also drive macroalgae communties indirectly through the alteration in
affect of another environmental driver. For instance, increasing degrees
of exposure may positively influence the amount of area available to
trap light on macroalgal fronds, as well as increasing nutrient uptake
due to increased turbulence in the boundry layer around the frond
{[}@lobban1994{]}. The most important direct effect of wave exposure on
macroalgal communities is through mechanical dislodgment, which
ultimately leads to expiration. Wave exposure is a complex abiotic
variable which varies spatially and temporarly in the marine enviroment.
Furthermore, the degree to which a macroalgae community is exposed which
depend on local site characteristics, such as bathymetry and local wind
patterns. Despite this fact, macrolagae have been able to persist in
often harsh and variable wave environments. Macroalage are sessile
organisms and incapabable of migrating when local conditions become
unsuitable. Therefore, macroalgae must adapt to the local wave climate
in order to persist and survive, and achieve this through morphological
variation.

\begin{quote}
keeping this for now @moss1948 investigated the anatomy, chemical
composistion of \emph{Fucus spiralis} at three sites that varied in wave
exposure (sheltered, medium exposure and exposed). The authors found
that individuals in exposed sites showed less branching of thalli as
well as variation in organic nitrogen, mannitol, laminarian and alginic
acid concentrations.
\end{quote}
